Bridging the foundational gap: updating algebraic geometry in face of current challenges regarding formalizability, constructivity and predicativity

### 2021-11-20 slides-dagstuhl2021.tex
More slides for Dagstuhl

\usepackage{pgfpages}
%\setbeameroption{show notes on second screen}
%\setbeamertemplate{note page}[plain]
%\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\justifying\footnotesize#2\par}}}
\newcommand{\jnote}[2]{}

@@

  \bigskip
  %\includegraphics[height=0.32\textwidth]{olivia-lattices}
  \bigskip

@@

  \scriptsize
  \textit{-- an invitation --}
  \bigskip

@@

\begin{frame}{Algebraic geometry in a nutshell}
  Turn \hil{commutative rings} into \hil{spaces}, and \hil{glue} those
  spaces.
  \bigskip

@@

  \subhead{Concrete results}
  Fermat's Last Theorem: For~$n \geq 3$, no positive integers satisfy
  \[ a^n + b^n = c^n. \]
\end{frame}

\begin{frame}{Transfinite methods}
  The current presentation of algebraic geometry employs:
  \begin{itemize}
    \item large structures: classes, large categories, universes, \ldots
    \item law of excluded middle
    \item axiom of choice
  \end{itemize}
  \bigskip

  \subhead{despite:}
  \begin{enumerate}
    \item subject matter (in part) very concrete
    \item computer algebra systems for computations exist
    \item constructive algebra well-developed
  \end{enumerate}
\end{frame}

>>>>

\usepackage{pgfpages}
\usepackage[export]{adjustbox}
%\setbeameroption{show notes on second screen}
%\setbeamertemplate{note page}[plain]
%\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\justifying\footnotesize#2\par}}}
\newcommand{\jnote}[2]{}

@@

  \bigskip
  %\includegraphics[height=0.32\textwidth]{olivia-lattices}
  \bigskip
  \bigskip
  \bigskip
  \bigskip

@@

  \scriptsize
  \textit{-- an invitation --}
  \bigskip
  \bigskip
  \bigskip

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1}\end{minipage}}
\begin{frame}{Algebraic geometry in a nutshell}
  Turn \hil{commutative rings} into \hil{spaces}, and \hil{glue} those
  spaces.
  \bigskip

@@

  \subhead{Concrete results}
  Fermat's Last Theorem: For~$n \geq 3$, no positive integers satisfy
  \[ a^n + b^n = c^n. \]
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.95cm}\includegraphics[width=\paperwidth]{staircase}\end{minipage}}
\begin{frame}{Transfinite methods}
  The current presentation of algebraic geometry hinges on:
  \begin{itemize}
    \item large structures: classes, large categories, universes, \ldots
    \item powersets
    \item law of excluded middle
    \item axiom of choice
  \end{itemize}
  \bigskip

  \subhead{despite:}
  \begin{enumerate}
    \item subject matter (in part) very concrete
    \item practical computer algebra systems for computations
    \item well-developed constructive algebra
  \end{enumerate}
\end{frame}}

\begin{frame}{Formalizing algebraic geometry}
  \bigskip
  \bigskip
  \bigskip
  \includegraphics[width=\textwidth]{schemes-in-lean2}

  \vspace*{-17em}
  \includegraphics[frame=0.1em,bb=0 170 610 625,width=0.3\textwidth]{schemes-in-coq}\hfill
  \includegraphics[frame=0.1em,bb=0 170 610 625,width=0.3\textwidth]{schemes-in-lean}
\end{frame}

\begin{frame}{A trinitarian challenge}
  \begin{center}\tiny $1^3 + \cdots + n^3 = (1 + \cdots + n)^2$\end{center}
  \vspace*{-0.6em}

  \begin{columns}
    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{lem}

      \hil{constructivity}
      \par
    \end{column}
    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{sum-of-cubics}

      \hil{elegance}
      \par
    \end{column}

    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{proof-assistants}

      \par
      \hil{formalizability}
    \end{column}
  \end{columns}
  \bigskip
  \bigskip

  \subhead{Thesis}
  \centering
  Elegant mathematics lends itself to rewarding formalization.
\end{frame}
